---
title: "Untitled"
author: "Juan Carlos Villaseñor-Derbez"
date: "last update `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
  library(sparklyr)
  library(magrittr)
  library(janitor)
  library(taxize)
  library(tmap)
  library(sf)
  library(modelr)
  library(caret)
  library(recipes)
  library(yardstick)
  library(tidyposterior)
  library(rsample)
  library(tidyverse)
})

get_mode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

# Test 2

The previous data exploration was a mess, and I need to re-start wrangling the data to have a dataset where I can actually ask many different questions.

Some of the questions are:

- Are `fishing hours` (from GFW) a better proxy of fishing effort than of `fishing days` (from Chilean data)?
- Can these (along with many boatl-level) be used to predict catches?
- How can we assign catches (reported on land) to individual points on the map?

## Load data

### Chile landings data

```{r}
chile_catches <- read.csv(here::here("raw_data","chile_landings_with_GFW.csv"), stringsAsFactors = F, na.strings = "null") %>% 
  clean_names() %>% 
  filter(matched,
         species == "ANCHOVETA",
         fishing_gear == "CERCO") %>% 
  mutate(port_name = taxize_capwords(port_name, strict = T, onlyfirst = T),
         species = taxize_capwords(species, strict = T, onlyfirst = T),
         fishing_gear = taxize_capwords(fishing_gear, strict = T, onlyfirst = T),
         v_name = taxize_capwords(v_name, strict = T, onlyfirst = T),
         departure = lubridate::month(departure),
         arrival = lubridate::month(arrival),
         distance_traveled = as.numeric(distance_traveled),
         max_distance_from_shore = as.numeric(max_distance_from_shore),
         power = as.numeric(power),
         aux_power = as.numeric(aux_power)) %>% 
  group_by(trip_id, year, departure, arrival, length, capacity, power, aux_power, trip_days, distance_traveled, max_distance_from_shore) %>% 
  summarize(landings = sum(ton, na.rm = T)) %>% 
  ungroup() %>% 
  filter(landings > 1)
```

This dataset was provided by Renato, although Juan had already played with it. The file contains `r ncol(chile_catches)` variables, which include year-boat-trip-species landings data. We can identify a single trip by the boat name, the day they left port and when they returned. There are a total of `r length(unique(chile_catches$v_name))` across `r length(unique(chile_catches$year))` years. Juan was able to match some of these boats (`r length(unique(chile_catches$v_name[chile_catches$matched]))` out of `r length(unique(chile_catches$v_name))`). From these matched vessels, we have a total of `r length(unique(chile_catches$trip_id[chile_catches$matched]))` trips made, from which we can derive information.

### GFW data

We want to add to this data some information from GFW. This dataset contains fishing hours for each position estimated from VMS data, as well as an indication on wheather the boat was fishing or not.

```{r}
gfw_data <- readRDS(here::here("raw_data","chilean_fleet_w_gfw.rds"))
```

```{r}
trip_info <- gfw_data %>% 
  filter(nnet_score == 1,
         between(lon, -90, -65),
         between(lat, -60, 10)) %>%
  group_by(mmsi, trip_id) %>% 
  summarize(min_lat = min(lat),
            max_lat = max(lat),
            mean_lat = mean(lat),
            median_lat = median(lat),
            mode_lat = get_mode(lat),
            min_lon = min(lon),
            max_lon = max(lon),
            mean_lon = mean(lon),
            median_lon = median(lon),
            mode_lon = get_mode(lon),
            hours = sum(hours)) %>% 
  ungroup() %>%
  select(trip_id, mmsi, everything())

chile_catches_full <- chile_catches %>%  
  left_join(trip_info, by = "trip_id") %>% 
  filter(!is.na(mmsi))
```


```{r}
set.seed(43)
chile_split <- rsample::initial_split(chile_catches_full, prop = 0.8, strata = "year")

chile_train <- rsample::training(chile_split)

chile_test  <- rsample::testing(chile_split)
```

```{r}
chile_train %>% 
  mutate(source = "train") %>% 
  rbind(chile_test %>% mutate(source = "test")) %>% 
  group_by(year, source) %>% 
  count() %>% 
  group_by(source) %>% 
  mutate(n = n/sum(n)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = n, fill = source)) +
  geom_col(position = "dodge")
```


```{r}
chile_vfold <- rsample::vfold_cv(chile_train, v = 5, repeats = 1, strata = "year")
```

```{r}
lm_foo <- function(dep_var, ind_vars, split) {
  
  formula <- as.formula(paste0(dep_var, "~", paste(ind_vars, collapse = "+")))
  
  fitted_lm <- lm(formula, data = rsample::analysis(split))
  
  lm_performance <- rsample::assessment(split) %>%
    modelr::add_predictions(model = fitted_lm)
  
  lm_rq <- yardstick::rsq(lm_performance, truth = landings, estimate = pred)
}
```

```{r}
dep_var <- "landings"
ind_vars <- c(".")

chile_vfold %>% 
  mutate(rsquare = map_dbl(splits, lm_foo, dep_var = dep_var, ind_vars = ind_vars)) %>% 
  ggplot(aes(rsquare)) + 
  geom_histogram()
```


```{r}
candidate_variables <- chile_catches_full %>% 
  select(-c(landings, mmsi, trip_id, departure, arrival, year)) %>% 
  colnames()

third_rule <- round(length(candidate_variables) / 3)
  
candidate_params <-expand.grid(mtry = c(2, third_rule, length(candidate_variables)),
                               splitrule = c("variance", "extratrees", "maxstat")) %>%
  as_data_frame()

candidate_params
```

```{r}
chile_recipe <- chile_catches_full %>%
  select(dep_var, candidate_variables) %>% 
  recipes::recipe(landings ~ ., data = .) %>% 
  recipes::step_log(all_outcomes()) %>% 
  recipes::step_center(all_predictors()) %>% 
  recipes::step_scale(all_predictors()) %>% 
  recipes::step_meanimpute(all_predictors())

chile_recipe
```

```{r}
chile_prepped <- recipes::prep(chile_recipe, training = chile_catches_full)

chile_prepped
```


```{r}
chile_baked <- recipes::bake(chile_prepped, newdata = chile_catches_full)

head(chile_baked)
```

```{r}
fit_forests <-   function(mtry, splitrule, splits, candidate_variables, my_recipe){
  
  test_ranger <- function(split, candidate_variables, mtry, splitrule, my_recipe) {
    
    analysis_data <- rsample::analysis(split)
    
    prepped_recipe <-recipes::prep(my_recipe, training = analysis_data)
    
    analysis_data <- bake(prepped_recipe, newdata = analysis_data)
    
    assessment_data <- bake(prepped_recipe, newdata = rsample::assessment(split))
  
    fitted_forest <- ranger::ranger(landings ~ ., data = analysis_data, mtry = mtry, splitrule = splitrule)
    
    assessment_data$prediction = predict(fitted_forest, data = assessment_data)$predictions
    
    metric <- yardstick::rmse(assessment_data, truth = landings, estimate = prediction)
  }
  
  splits <- splits %>%
    mutate(fitted_forest = map_dbl(splits,
                                   test_ranger,
                                   candidate_variables = candidate_variables,
                                   mtry = mtry,
                                   splitrule = splitrule,
                                   my_recipe = my_recipe))
  
  mean_metric <- mean(splits$fitted_forest)
}
```


```{r}
candidate_params <- candidate_params %>%
  mutate(ranger_rmse = map2_dbl(mtry,
                                splitrule,
                                fit_forests,
                                splits = chile_vfold,
                                candidate_variables = candidate_variables,
                                my_recipe = chile_recipe))
```


```{r}
candidate_params <- candidate_params %>%
  arrange(ranger_rmse)

best_mtry <- candidate_params$mtry[1]

best_splitrule <- candidate_params$splitrule[1]

candidate_params %>%
  ggplot(aes(splitrule, ranger_rmse, fill = factor(mtry))) +
  geom_col(position = "dodge")
```

```{r}
lm_foo <- function(split, ind_vars, my_recipe) {
  formula <- as.formula(paste0("landings ~ ", paste(ind_vars, collapse = "+")))
  
  prepped_recipe <- recipes::prep(my_recipe, training = rsample::analysis(split))
  
  analysis_data <- bake(prepped_recipe, newdata = rsample::analysis(split))
  
  assessment_data <- bake(prepped_recipe, newdata = rsample::assessment(split))
  
  fitted_lm <- lm(landings ~ ., data = analysis_data)
  
  lm_performance <- assessment_data %>%
  modelr::add_predictions(model = fitted_lm)
  
  metric <- yardstick::rmse(lm_performance, truth = landings, estimate = pred)
}

ranger_foo <- function(split, candidate_variables, dep_var, mtry, splitrule, my_recipe) {
  analysis_data <- rsample::analysis(split)
  
  prepped_recipe <- recipes::prep(my_recipe, training = analysis_data)
  
  analysis_data <- bake(prepped_recipe, newdata = analysis_data)
  
  assessment_data <- bake(prepped_recipe, newdata = rsample::assessment(split))

  fitted_forest <- ranger::ranger(landings ~ .,
                                  data = analysis_data,
                                  mtry = mtry,
                                  splitrule = splitrule)
  
  assessment_data$prediction = predict(fitted_forest, data = assessment_data)$predictions
  
  metric <- yardstick::rmse(assessment_data, truth = landings, estimate = prediction)
}

chile_vfold <- chile_vfold %>%
  mutate(lm_rmse = map_dbl(splits,
                           lm_foo,
                           ind_vars = candidate_variables,
                           my_recipe = chile_recipe),
         ranger_rmse = map_dbl(splits,
                               ranger_foo,
                               candidate_variables = candidate_variables,
                               mtry = best_mtry,
                               splitrule = best_splitrule,
                               my_recipe = chile_recipe))

chile_vfold
```

```{r}
chile_performance <- chile_vfold %>%
  perf_mod(seed = 42)

forest_v_lm <- contrast_models(chile_performance, "ranger_rmse", "lm_rmse")

summary(forest_v_lm)
```

```{r}
ggplot(forest_v_lm, size = 0.1)
```

```{r}
chile_prepped <- prep(chile_recipe, training = chile_train)

fitted_forest <- ranger::ranger(
  landings ~ .,
  data =  bake(chile_prepped, newdata = chile_train),
  mtry = best_mtry,
  splitrule = best_splitrule,
  importance = "permutation")

fitted_lm <- lm(landings ~ ., data = bake(chile_prepped, newdata = chile_train))
```

```{r}
data_frame(variable = names(ranger::importance(fitted_forest)),
           importance = ranger::importance(fitted_forest)) %>%
  mutate(variable = fct_reorder(variable, importance)) %>%
  ggplot(aes(variable, importance)) +
  geom_col() +
  coord_flip()
```

Compare fit on training data

```{r}
chile_results <- bake(chile_prepped, newdata = chile_train) %>%
  mutate(lm_pred = predict(fitted_lm, newdata = bake(chile_prepped, newdata = chile_train)),
         forest_pred = predict(fitted_forest, data = bake(chile_prepped, newdata = chile_train))$predictions)


chile_results %>% 
  select(landings, lm_pred, forest_pred) %>%
  gather(model, prediction, contains("pred")) %>% 
  ggplot(aes(landings, prediction, color = model)) +
  geom_point() +
  geom_abline(aes(intercept = 0, slope = 1), color = "red") +
  facet_wrap( ~ model) +
  cowplot::theme_cowplot()
```

```{r}
cor(chile_results$landings, chile_results$forest_pred)^2
cor(chile_results$landings, chile_results$lm_pred)^2
```

Compare fit on testing data

```{r}
chile_results <- bake(chile_prepped, newdata = chile_test) %>%
  mutate(lm_pred = predict(fitted_lm, newdata = bake(chile_prepped, newdata = chile_test)),
         forest_pred = predict(fitted_forest, data = bake(chile_prepped, newdata = chile_test))$predictions)


chile_results %>% 
  select(landings, lm_pred, forest_pred) %>%
  gather(model, prediction, contains("pred")) %>% 
  ggplot(aes(landings, prediction, color = model)) +
  geom_point() +
  geom_abline(aes(intercept = 0, slope = 1), color = "red") +
  facet_wrap( ~ model) +
  cowplot::theme_cowplot()
```

```{r}
cor(chile_results$landings, chile_results$forest_pred)^2
cor(chile_results$landings, chile_results$lm_pred)^2
```


```{r}
fit_control <- trainControl(method = "repeatedcv",
                            number = 5,
                            repeats = 1,
                            allowParallel = TRUE)

caret_forest <- caret::train(chile_recipe,
                             data = chile_train,
                             method = "ranger",
                             trControl = fit_control)


chile_results <- chile_results %>% 
  mutate(caret_pred = predict(caret_forest$finalModel,
                              data = select(chile_results, candidate_variables))$predictions)

chile_results %>% 
  select(landings, lm_pred, forest_pred, caret_pred) %>%
  gather(model, prediction, contains("pred")) %>%
  ggplot(aes(landings, prediction, color = model)) +
  geom_point() +
  geom_abline(aes(intercept = 0, slope = 1), color = "red") +
  facet_wrap( ~ model) +
  cowplot::theme_cowplot()
```


```{r}
cor(chile_results$landings, chile_results$forest_pred)
cor(chile_results$landings, chile_results$lm_pred)
cor(chile_results$landings, chile_results$caret_pred)
```

# Next stepps

1 Improve model fit by adding SST + Chl-a + lagrangian structures?
2 Think of ways to include "areas" / "regions" other than min/max/mode/mean of Lat / Long
3 Include variance of the above
4 Apply a similar method to SSt + Chla + lagrangian
5 Use other algorithms
6 Two parts:
  - Using GFW to predict catches of in-sample boats
  - Can we use little data we have on this dataset to recreate timeseries of catches? (i.e. compare this to FAO data / SAU / official landings)










